<!--loading necessary packages-->
```{r loading packages, include=FALSE, warning= FALSE}
#unsere packages
library(tidyverse)
library(readxl)
library(tinytex)
library(broom)
library(viridis) # color-schemes
library(knitr)
library(kableExtra)
library(stargazer) #Darstellung Tests/Regressionen
library(ggpubr) #Darstellung von mehreren plots in einer Abbildung

```

## Our Dataset


- we look at the survey data provided by the Survey of Consumer Expectations of the New York FED
- Q8v2,Q8v2part2,Q9 are relevant for the expectations

D’Acunto (2022):
- explains Survey of Consumer Expectations (SCE)
- shortcoming: only a small time series because not that old. "anchors" answers with their questions
- but overall good material
- disagreements increase volatility seen in covid19 2020 data
- looking at IQR
- only in the covid19 crisis short/long-term expectations differ


<!--short term Expectations by the SCE-->
```{r, include=FALSE, warning= FALSE}
#load data
SCE_Data <- read_excel("DATA/FRBNY-SCE-Data.xlsx", sheet = "Inflation expectations", skip = 2)

#rename columns, change format of the date column
SCE_Data <- SCE_Data %>%
  rename(Median=`Median one-year ahead expected inflation rate`,Lower_quantil=`25th Percentile one-year ahead expected inflation rate`, Upper_quantil=`75th Percentile one-year ahead expected inflation rate`,Point_prediction=`Median point prediction one-year ahead inflation rate`)%>%
  mutate(Date=...1*100+1)%>%
  select(c(2,4,6,8,10))%>%
  relocate(Date, .before = Median)

SCE_Data$Date <- as.Date(as.character(SCE_Data$Date),format = "%Y%m%d")



```

<!--Uncertainty data-->
```{r,include=FALSE, warning= FALSE}
SCE_Data_unc <- read_excel("DATA/FRBNY-SCE-Data.xlsx", sheet = "Inflation uncertainty", skip = 2)

SCE_Data_unc <- SCE_Data_unc%>%
  rename(Uncertainty=`Median one-year ahead uncertainty`)


```



<!--load Demographics-->
```{r, include=FALSE, warning= FALSE}
#load data from demographic sheet
SCE_Data_demo <- read_excel("DATA/FRBNY-SCE-Data.xlsx", sheet = "Inflation expectations Demo", skip = 3)

```


<!--CPI Data-->
```{r, include=FALSE, warning= FALSE}
#load CPI
cpi <- read_csv("DATA/CPIAUCSL.csv")

  
SCE_Data_summary <-bind_cols(SCE_Data, SCE_Data_unc[2],SCE_Data_demo[2:7],cpi[2])%>%
  rename("HS or Less"=`Education High School or Less`,"College"=`Education Some College`,"BA or Higher"=`Education BA or Higher`, CPI=CPIAUCSL_PC1)

```



- a plot of consumer expectations for one-year expectations of the last ten years give a first expression



```{r,echo=FALSE, warning=FALSE, fig.cap = "Short-term Expectations from 2013-present"}

#pivot longer
SCE_plot <- select(SCE_Data_summary,c(1:5))%>%
  pivot_longer(cols=c(2:5), names_to = "Measures", values_to = "Expectations", values_drop_na = FALSE)


#reihenfolge der quantile
SCE_plot$Measures <- factor(SCE_plot$Measures, levels = c("Lower_quantil","Median","Point_prediction","Upper_quantil"))


ggplot(data=SCE_plot) + aes(x=Date, y = Expectations, color = Measures) + geom_line(size=0.8)+scale_color_viridis(discrete = TRUE)+labs(x="Years", y="Monthly Expectations")+scale_x_date(date_labels = "%Y",date_breaks = "2 years")+ theme_minimal()



```

- we see different pattern of behavior which allwos us to split the data in two time intervals.
- we will analyze them in the next chapter descriptive statistics

\newpage


## Descriptive statistics

- we now look more into detail of the inflation expectations dataset
- we  also look at uncertainty, describe it


<!--Descriptive summaries-->
```{r, include=FALSE, warning= FALSE}
#sub-data sets for both intervals

SCE_Data_01 <- SCE_Data_summary%>%
  filter(Date<"2020-05-01")

SCE_Data_02 <- SCE_Data_summary%>%
  filter(Date>"2020-04-01")


#summary for first interval 13-20
sum_01 <- SCE_Data_01%>%
  summarise(Mean = mean(Median),
            Volatility = sd(Median))%>%
  mutate(Measure="Median")%>%
  relocate(Measure, .before = Mean)

sum_02 <- SCE_Data_01%>%
  summarise(Mean = mean(Point_prediction),
            Volatility = sd(Point_prediction))%>%
  mutate(Measure="Point prediction")%>%
  relocate(Measure, .before = Mean)

sum_03 <- SCE_Data_01%>%
  summarise(Mean = mean(Uncertainty),
            Volatility = sd(Uncertainty))%>%
  mutate(Measure="Uncertainty")%>%
  relocate(Measure, .before = Mean)

sum_04 <- SCE_Data_01%>%
  summarise(Mean = mean(CPI),
            Volatility = sd(CPI))%>%
  mutate(Measure="CPI")%>%
  relocate(Measure, .before = Mean)


#summary for second interval

sum_05 <- SCE_Data_02%>%
  summarise(Mean = mean(Median),
            Volatility = sd(Median))%>%
  mutate(Measure="Median")%>%
  relocate(Measure, .before = Mean)

sum_06 <- SCE_Data_02%>%
  summarise(Mean = mean(Point_prediction),
            Volatility = sd(Point_prediction))%>%
  mutate(Measure="Point prediction")%>%
  relocate(Measure, .before = Mean)

sum_07 <- SCE_Data_02%>%
  summarise(Mean = mean(Uncertainty),
            Volatility = sd(Uncertainty))%>%
  mutate(Measure="Uncertainty")%>%
  relocate(Measure, .before = Mean)

sum_08 <- SCE_Data_02%>%
  summarise(Mean = mean(CPI),
            Volatility = sd(CPI))%>%
  mutate(Measure="CPI")%>%
  relocate(Measure, .before = Mean)


```

- we can see two different phases for inflation expectations: from June 2013 until April 2020 and from May 2020 until today
- first phase: stable expectations with median and point predictions at around 2.5-3.0% inflation
- second phase: sudden increase in expectations with a peak at arond 6.5% in the last months


```{r,echo=FALSE, warning=FALSE}
summary_01 <- bind_rows(sum_01,sum_02, sum_03)  


#darstellung mit kableExtra
  knitr::kable(summary_01,
  digits = 4,caption="Summary for expectations from 2013-2020")%>%
    kable_styling(latex_options = c("striped"))
```

```{r,echo=FALSE, warning=FALSE}
summary_02 <- bind_rows(sum_04,sum_05,sum_06)  


#darstellung mit kableExtra
  knitr::kable(summary_02,
  digits = 4,caption="Summary for expectations from 2020-present")%>%
    kable_styling(latex_options = c("striped"))
```
- we see higher values for average median, point estimate, volatility and uncertainty and therefore a wider range in the second sample


\newpage


## Some Demographic Differences

-we look at two major differences at the demographic level of our data set, namely sex and age

<!--Distribution function Age-->
```{r, echo=FALSE, warning=FALSE, fig.cap = "Distribution function for Age Groups"}
#show distribution for Age and Education

SCE_Data_age <- SCE_Data_summary%>%
  select(c(1,7:9))%>%
  pivot_longer(cols=c(2:4), names_to = "Age Groups", values_to = "Expectations", values_drop_na = FALSE)

#Rang of Age Groups
SCE_Data_age$`Age Groups` <- factor(SCE_Data_age$`Age Groups`, levels = c("Age Under 40","Age 40-60","Age Over 60"))


ggplot(data=SCE_Data_age, aes(x=Expectations, group=`Age Groups`, fill=`Age Groups`)) + geom_density(adjust=0.8, alpha=0.75) + scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) + labs(x="Inflation Expectations",y="Distribution")+scale_x_continuous(breaks = seq(0, 8, by = 1))+ theme_minimal()

```


<!--Distribution function Education-->
```{r, echo=FALSE, warning=FALSE, fig.cap = "Distribution function for Education Groups"}
#show distribution for Education

SCE_Data_ed <- SCE_Data_summary%>%
  select(c(1,10:12))%>%
  pivot_longer(cols=c(2:4), names_to = "Education Levels", values_to = "Expectations", values_drop_na = FALSE)

#Rang of Education Levels
SCE_Data_ed$`Education Levels` <- factor(SCE_Data_ed$`Education Levels`, levels = c("HS or Less","College","BA or Higher"))


ggplot(data=SCE_Data_ed, aes(x=Expectations, group=`Education Levels`, fill=`Education Levels`)) + geom_density(adjust=0.8, alpha=0.75) + scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) + labs(x="Inflation Expectations",y="Distribution")+scale_x_continuous(breaks = seq(0, 7, by = 1))+ theme_minimal()

```


- the next dataset contains the google searches for inflation or inflation rate in the US

<!--Trends data for USA-->
```{r, include=FALSE, warning= FALSE}
#loading the data
infl_US <- read_delim("DATA/infl_US.csv",delim = ";", escape_double = FALSE, col_types = cols(Monat = col_date(format = "%d/%m/%Y")), trim_ws = TRUE)

infl_r_US <- read_delim("DATA/infl_r_US.csv",delim = ";", escape_double = FALSE, col_types = cols(Monat = col_date(format = "%d/%m/%Y")), trim_ws = TRUE)

```


<!--Google trends for USA-->
```{r, echo=FALSE, warning=FALSE, fig.cap = "US Google Trends Results for Inflation/Inflation rate from 2010-present"}

infl_summary <- bind_cols(infl_US,infl_r_US)%>%
  select(c(1,2,4,))%>%
  rename(Date=Monat...1,Inflation=`inflation: (Vereinigte Staaten)`,"Inflation rate"=`inflation rate: (Vereinigte Staaten)`)

infl_plot <- infl_summary%>%
  pivot_longer(cols=2:3, names_to = "Searched Terms", values_to = "Relative Rankings", values_drop_na = FALSE)

ggplot(data=infl_plot) + aes(x=Date, y =`Relative Rankings`, color =`Searched Terms`) + geom_line(size=0.9,alpha=1)+scale_color_manual(values=c("#00798c","#B9314F"))+labs(x="Years")+scale_x_date(date_labels = "%Y",date_breaks = "1 year")+ theme_minimal()  


```

- google trends for the US show stable behavior until second half of 2021, after that huge increase
- inflation and inflation rate are very similar in their relative search frequency
- we will later combine these data with the SCE data set in some regression models




\newpage