<!--loading necessary packages-->
```{r loading packages, include=FALSE, warning= FALSE}
#unsere packages
library(tidyverse)
library(readxl)
library(tinytex)
library(broom)
library(viridis) # color-schemes
library(knitr)
library(kableExtra)
library(stargazer) #Darstellung Tests/Regressionen
library(ggpubr) #Darstellung von mehreren plots in einer Abbildung

```

## Our Dataset


- we look at the survey data provided by the Survey of Consumer Expectations of the New York FED and the MSC data from the survey of consumers

D’Acunto (2022):
- explains Survey of Consumer Expectations (SCE)
- shortcoming: only a small time series because not that old. "anchors" answers with their questions
- but overall good material
- disagreements increase volatility seen in covid19 2020 data
- looking at IQR
- only in the covid19 crisis short/long-term expectations differ


<!--short term Expectations by the SCE-->
```{r, include=FALSE, warning= FALSE}
#load data
SCE_Data <- read_excel("DATA/FRBNY-SCE-Data.xlsx", sheet = "Inflation expectations", skip = 2)

#rename columns, change format of the date column
SCE_Data <- SCE_Data %>%
  rename(median_1y=`Median one-year ahead expected inflation rate`,lower_quantil_1y=`25th Percentile one-year ahead expected inflation rate`, upper_quantil_1y=`75th Percentile one-year ahead expected inflation rate`,point_prediction_1y=`Median point prediction one-year ahead inflation rate`,median_3y=`Median three-year ahead expected inflation rate`,lower_quantil_3y=`25th Percentile three-year ahead expected inflation rate`, upper_quantil_3y=`75th Percentile three-year ahead expected inflation rate`,point_prediction_3y=`Median point prediction three-year ahead inflation rate`)%>%
  mutate(date=...1*100+1)%>%
  select(2:10)%>%
  relocate(date, .before = median_1y)

SCE_Data$date <- as.Date(as.character(SCE_Data$date),format = "%Y%m%d")



```

- two plots of consumer expectations for one-year and three-year expectations of the last ten years give a first expression



```{r,echo=FALSE, warning=FALSE, fig.cap = "Expectations from 2013-present"}

#pivot one-year and three-year
SCE_plot_1y <- select(SCE_Data,c(1,2,4,6,8))%>%
  pivot_longer(cols=c(2:5), names_to = "quantiles", values_to = "expectations", values_drop_na = FALSE)

SCE_plot_3y <- select(SCE_Data,c(1,3,5,7,9))%>%
  pivot_longer(cols=c(2:5), names_to = "quantiles", values_to = "expectations", values_drop_na = FALSE)

#reihenfolge der quantile
SCE_plot_1y$quantiles <- factor(SCE_plot_1y$quantiles, levels = c("lower_quantil_1y","median_1y","point_prediction_1y","upper_quantil_1y"))

SCE_plot_3y$quantiles <- factor(SCE_plot_3y$quantiles, levels = c("lower_quantil_3y","median_3y","point_prediction_3y","upper_quantil_3y"))


ggplot(data=SCE_plot_1y) + aes(x=date, y = expectations, color = quantiles) + geom_line(size=0.8)+scale_color_viridis(discrete = TRUE)+labs(x="years", y="monthly expectations")+scale_x_date(date_labels = "%Y",date_breaks = "2 years")+ theme_minimal()

ggplot(data=SCE_plot_3y) + aes(x=date, y = expectations, color = quantiles) + geom_line(size=0.8)+scale_color_viridis(discrete = TRUE)+labs(x="years", y="monthly expectations")+scale_x_date(date_labels = "%Y",date_breaks = "2 years")+ theme_minimal()


```

- we see different pattern of behavior which allwos us to split the data in two time intervals.
- we will analyze them in the next chapter descriptive statistics
- the next two dataset contain the google trends for searches for inflation or inflation rate in Germany and the US

<!--Trends data for Germany/USA-->
```{r, include=FALSE, warning= FALSE}
#loading the data
infl_Ger <- read_delim("DATA/infl_Ger.csv",delim = ";", escape_double = FALSE, col_types = cols(Monat = col_date(format = "%d/%m/%Y")), trim_ws = TRUE)

infl_r_Ger <- read_delim("DATA/infl_r_Ger.csv",delim = ";", escape_double = FALSE, col_types = cols(Monat = col_date(format = "%d/%m/%Y")), trim_ws = TRUE)

infl_US <- read_delim("DATA/infl_US.csv",delim = ";", escape_double = FALSE, col_types = cols(Monat = col_date(format = "%d/%m/%Y")), trim_ws = TRUE)

infl_r_US <- read_delim("DATA/infl_r_US.csv",delim = ";", escape_double = FALSE, col_types = cols(Monat = col_date(format = "%d/%m/%Y")), trim_ws = TRUE)

```


<!--Google trends for Germany-->
```{r, echo=FALSE, warning=FALSE, fig.cap = "German Google Trends from 2010-present"}


infl_sum_Ger <- bind_cols(infl_Ger,infl_r_Ger)%>%
  select(c(1,2,4,))%>%
  rename(date=Monat...1,inflation=`inflation: (Deutschland)`,inflationsrate=`inflationsrate: (Deutschland)`)%>%
  pivot_longer(cols=2:3, names_to = "searched_terms", values_to = "relative_rankings", values_drop_na = FALSE)

ggplot(data=infl_sum_Ger) + aes(x=date, y = relative_rankings, color = searched_terms) + geom_line(size=0.9, alpha=1)+scale_color_manual(values=c("#00798c","#edae49"))+labs(x="years")+scale_x_date(date_labels = "%Y",date_breaks = "1 year")+ theme_minimal()  



```

- we see different searching behavior for the terms "inflation" and "inflationsrate" with the highest relative shares in the last months


<!--Google trends for USA-->
```{r, echo=FALSE, warning=FALSE, fig.cap = "US Google Trends from 2010-present"}

infl_sum_US <- bind_cols(infl_US,infl_r_US)%>%
  select(c(1,2,4,))%>%
  rename(date=Monat...1,inflation=`inflation: (Vereinigte Staaten)`,inflation_rate=`inflation rate: (Vereinigte Staaten)`)%>%
  pivot_longer(cols=2:3, names_to = "searched_terms", values_to = "relative_rankings", values_drop_na = FALSE)

ggplot(data=infl_sum_US) + aes(x=date, y = relative_rankings,color = searched_terms) + geom_line(size=0.9,alpha=1)+scale_color_manual(values=c("#00798c","#edae49"))+labs(x="years")+scale_x_date(date_labels = "%Y",date_breaks = "1 year")+ theme_minimal()  


```

- google trends for the US show a similar, but different pattern
- inflation and inflation rate are much more similar in their relative search frequency. but like in germany the highest rates are in the last month

\newpage


## Descriptive statistics

- we now look more into detail of the inflation expectations dataset

<!--Descriptive statistics-->
```{r, include=FALSE, warning= FALSE}
#sub-data sets for both intervals

CE_Data_01 <- SCE_Data%>%
  filter(date<"2020-05-01")

CE_Data_02 <- SCE_Data%>%
  filter(date>"2020-04-01")


#summary for first interval 13-20
sum_01 <- CE_Data_01%>%
  summarise(Mean = mean(median_1y),
            Volatility = sd(median_1y))%>%
  mutate(Measure="Median")%>%
  relocate(Measure, .before = Mean)

sum_02 <- CE_Data_01%>%
  summarise(Mean = mean(point_prediction_1y),
            Volatility = sd(point_prediction_1y))%>%
  mutate(Measure="Point prediction")%>%
  relocate(Measure, .before = Mean)

#summary for second interval

sum_03 <- CE_Data_02%>%
  summarise(Mean = mean(median_1y),
            Volatility = sd(median_1y))%>%
  mutate(Measure="Median")%>%
  relocate(Measure, .before = Mean)

sum_04 <- CE_Data_02%>%
  summarise(Mean = mean(point_prediction_1y),
            Volatility = sd(point_prediction_1y))%>%
  mutate(measure="Point prediction")%>%
  relocate(measure, .before = Mean)


```

- we can see two different phases for inflation expectations: from June 2013 until April 2020 and from May 2020 until today
- first phase: stable expectations with median and point predictions at around 2.5-3.0% inflation
- second phase: sudden increase in expectations with a peak at arond 6.5% in the last months

- we can see the differences in the descriptive statistics and growth rates for the point estimate of both intervals in the following tables



```{r,echo=FALSE, warning=FALSE}
summary_01 <- bind_rows(sum_01,sum_02)  


#darstellung mit kableExtra
  knitr::kable(summary_01,
  digits = 4,caption="Summary for expectations from 2013-2020")%>%
    kable_styling(latex_options = c("striped"))
```

```{r,echo=FALSE, warning=FALSE}
summary_02 <- bind_rows(sum_03,sum_04)  


#darstellung mit kableExtra
  knitr::kable(summary_02,
  digits = 4,caption="Summary for expectations from 2020-present")%>%
    kable_styling(latex_options = c("striped"))
```
- we see higher value for average median, point estimate, volatility and therefore a wider range in the second sample

- we can also see higher movements by comparing the growth rates of both intervals


<!--growth rates for first interval-->
```{r,echo=FALSE, warning=FALSE, fig.cap = "Growth rates of median expectations from 2013-20"}

#growth rate with index of beginning of sample

CE_Data_01 <- CE_Data_01%>%
  mutate(growth_rate=100*(median_1y-median_1y[1])/median_1y[1])

ggplot(data=CE_Data_01) + aes(x=date, y = growth_rate) + geom_line(size=1.2,color = "#00798c")+labs(x="years", y="growth rate to basis year")+scale_x_date(date_labels = "%Y",date_breaks = "1 year")+ theme_minimal()

```


-new picture now





- only small movements, decrease of about 15% in seven years

<!--growth rates-->
```{r, echo=FALSE, warning=FALSE, fig.cap = "Growth rates of median expectations from 2020-present"}

#growth rate with index of beginning of sample
CE_Data_02 <- CE_Data_02%>%
  mutate(growth_rate=100*(median_1y-median_1y[1])/median_1y[1])

#plot
ggplot(data=CE_Data_02) + aes(x=date, y = growth_rate) + geom_line(size=1.2,color = "#00798c")+labs(x="years", y="growth rate to basis year")+scale_x_date(date_labels = "%m-%Y",date_breaks = "3 months")+ theme_minimal()



```


- more than double the expectations in only two years! shows drastic shift in expectations


\newpage


## More statistics

-cyclical behavior of trends for US is very interesting
- no one looks after inflation in the summer, so maybe trends coincide with the sales in the winter (Black Friday)

<!--Google trends for USA-->
```{r, echo=FALSE, warning=FALSE, fig.cap = "US Google Trends from 2013-2016"}

infl_sum_US_sub <- bind_cols(infl_US,infl_r_US)%>%
  select(c(1,2,4,))%>%
  rename(date=Monat...1,inflation=`inflation: (Vereinigte Staaten)`,inflation_rate=`inflation rate: (Vereinigte Staaten)`)%>%
  pivot_longer(cols=2:3, names_to = "searched_terms", values_to = "relative_rankings", values_drop_na = FALSE)%>%
  filter(date>"2013-01-01")%>%
  filter(date<"2016-01-01")

ggplot(data=infl_sum_US_sub) + aes(x=date, y = relative_rankings,color = searched_terms) + geom_line(size=0.9,alpha=1)+scale_color_manual(values=c("#00798c","#edae49"))+labs(x="months")+scale_x_date(date_labels = "%m",date_breaks = "2 months")+ theme_minimal()  


```







\newpage