---
output:
  pdf_document: default
  html_document: default
---

## Updating Behavior

- to analyze updating behavior and its relationship with google searches or inflation expectations we need to compute the updating behavior of individuals at the micro level of the SCE data set
- we use the idea of dräger (ZITAT)



## Regressions

- we look now at the updating behavior
- analyzing this needs the micro data from the SCE



<!--Micro data by the SCE-->
```{r, include=FALSE, warning= FALSE}

#load all data
SCE_micro_13 <- read_excel("DATA/FRBNY-SCE-Public-Microdata-Complete-13-16.xlsx")
SCE_micro_17 <- read_excel("DATA/FRBNY-SCE-Public-Microdata-Complete-17-19.xlsx")
SCE_micro_20 <- read_excel("DATA/frbny-sce-public-microdata-latest.xlsx")

#combine in one
SCE_micro <- bind_rows(SCE_micro_13,SCE_micro_17,SCE_micro_20)


#tidy up
SCE_micro <- SCE_micro %>%
  rename(Expectation=Q8v2part2)%>%
  mutate(Date=date*100+1)%>%
  select(2:5)%>%
  relocate(Date, .before = userid)%>%
  group_by(userid)%>%
  mutate(last_exp=lag(Expectation))%>% 
  mutate(last_exp = coalesce(last_exp,Expectation))%>%
  mutate(dummy=case_when(Expectation-last_exp==0
 ~ 0,Expectation-last_exp!=0 ~ 1))%>%
  ungroup(userid)

SCE_micro$Date <- as.Date(as.character(SCE_micro$Date),format = "%Y%m%d")
SCE_micro[is.na(SCE_micro)] <- 0

SCE_micro_1 <- SCE_micro%>%
  count(Date)

SCE_micro_2 <- SCE_micro%>%
  group_by(Date)%>%
  summarise(monthly_count = sum(dummy))

SCE_micro_summary <-bind_cols(SCE_micro_1, SCE_micro_2[2])%>%
  filter(monthly_count>0)%>%
  mutate("Updating share"=monthly_count/n)
```



<!--Plot Micro data by the SCE all in one-->
```{r, include=FALSE, warning= FALSE}


ggplot(data=SCE_micro_summary) + aes(x=Date, y = `Updating share`) + geom_line(size=0.8,color="steelblue")+labs(x="Years",y="Updating Share")+scale_x_date(date_labels = "%Y",date_breaks = "year")+ theme_minimal()

```

-due to the shorter time series for the micro data we build two different regression tables with them we perform the different regressions


<!--Regressions-->
```{r, include=FALSE, warning= FALSE}
#transfer data into big regression summary, but without updating behavior
SCE_Data_regressions <-bind_cols(SCE_Data_summary,infl_summary[42:148,2:3])%>%
  rename(GS_infl=Inflation, GS_infl_r=`Inflation rate`)

#second regression table for updating and Google searches
SCE_Data_regressions_micro <-bind_cols(SCE_Data_summary[2:98,],SCE_micro_summary[4],infl_summary[43:139,2:3])%>%rename(GS_infl=Inflation, GS_infl_r=`Inflation rate`)

```


-first regression model with aggregated data: uncertainty~ GS_infl+GS_infl_r
-second regression model with aggregated data: Median~ GS_infl+GS_infl_r
- third regression model with aggregated data: median of exp~ GS_infl+GS_infl_r+Uncertainty+CPI

<!--Regression Model aggregated data-->
```{r, echo=FALSE, warning=FALSE, results='asis'}

#regression models
attach(SCE_Data_regressions)
model1 <- lm(Uncertainty~GS_infl+GS_infl_r)
model2 <- lm(Median~ GS_infl+GS_infl_r)
model3 <- lm(Median~ GS_infl+GS_infl_r+Uncertainty+CPI)
detach(SCE_Data_regressions)

#results for full sample aggr. data
stargazer(model1,model2,model3, header=FALSE, type='latex',model.numbers= FALSE,title = "Regressions for aggregated Data, full sample",omit.stat = c("rsq", "f"))


```


- we also differ between the two sub time-series to analyze if the correlations differ in the two phases

<!--Regression Model aggregated data sub-->
```{r, echo=FALSE, warning=FALSE}

#building sub sets
SCE_Data_regressions_01 <- SCE_Data_regressions%>%
  filter(Date<"2020-05-01")

SCE_Data_regressions_02 <- SCE_Data_regressions%>%
  filter(Date>"2020-04-01")

```

-we build two models for the first interval to analyze the effect of CPI

<!--Regression Models aggregated data sub1-->
```{r, echo=FALSE, warning=FALSE, results='asis'}

attach(SCE_Data_regressions_01)
model1_01 <- lm(Uncertainty~GS_infl+GS_infl_r)
model2_01 <- lm(Median~GS_infl+GS_infl_r)
model3_01 <- lm(Median~ GS_infl+GS_infl_r+Uncertainty+SCE_Data_regressions_01$CPI)
detach(SCE_Data_regressions_01)

stargazer(model1_01,model2_01,model3_01, header=FALSE, type='latex',model.numbers= FALSE,title = "Regressions for aggregated Data, sample from 06/13-04/20",omit.stat = c("rsq", "f"))

```

- we do the same regressions with our second sub time-series

<!--Regression Models aggregated data sub2-->
```{r, echo=FALSE, warning=FALSE, results='asis'}

attach(SCE_Data_regressions_02)
model1_02 <- lm(Uncertainty~GS_infl+GS_infl_r)
model2_02 <- lm(Median~GS_infl+GS_infl_r)
model3_02 <- lm(Median~ GS_infl+GS_infl_r+Uncertainty+SCE_Data_regressions_02$CPI)
detach(SCE_Data_regressions_02)

stargazer(model1_02,model2_02,model3_02, header=FALSE, type='latex',model.numbers= FALSE,title = "Regressions for aggregated Data, sample 05/20-04/22",omit.stat = c("rsq", "f"))




```


<!--Regression Model micro data 1-->
```{r, echo=FALSE, warning=FALSE, results='asis'}
#regressions for micro data
attach(SCE_Data_regressions_micro)
model1_micro <- lm(`Updating share`~ GS_infl+GS_infl_r)
model2_micro <- lm(Uncertainty~ `Updating share`)
model3_micro <- lm(Median~ `Updating share`)

stargazer(model1_micro,model2_micro,model3_micro, header=FALSE, type='latex',model.numbers= FALSE,title = "Regressions for Micro-Data",omit.stat = c("rsq", "f"))


```



